---
title: "Final Project"
author: "Maddox de Bretteville and Ryan Belkin"
date: "5/28/2023"
header-includes:
   - \usepackage{bbm, amsmath,amsfonts,amsthm,amssymb,mathrsfs,amsxtra,amscd,latexsym, xcolor, graphicx, fvextra}
   - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
output: pdf_document
---

\newcommand{\E}{\mathbb{E}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\Corr}{\mathrm{Corr}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\fix}[1]{\textcolor{red}{\textbf{\large (#1)\normalsize}}}

```{r}
#in order to run this code, you will need to install these packages. You can do so by going to the install button, or by going to the console and typing, for example, install.packages("knitr")
suppressPackageStartupMessages(library(knitr)) #makes pdfs
suppressPackageStartupMessages(library(latex2exp))
suppressPackageStartupMessages(library(ggplot2)) #makes nice plots
suppressPackageStartupMessages(library(tidyverse))
#good library for data manipulation, includes dplyr and ggplot
# you can read more about the tidyverse at: https://r4ds.had.co.nz/
knitr::opts_chunk$set(echo = TRUE)
```



library(dplyr)
NFL = read.csv("pbp-2022.csv")
NFL_one = filter(NFL, Down == 1, YardLine >= 90, Yards <=10)
NFL_pass = filter(NFL_one, PlayType == "PASS")
NFL_rush = filter(NFL, PlayType == "RUSH")
NFL_Chiefspass = filter(NFL_pass, OffenseTeam == "KC")
NFL_Chiefsrush = filter(NFL_rush, OffenseTeam == "KC")
NFL_Steelerspass = filter(NFL_pass, OffenseTeam == "PIT")
NFL_Steelersrush = filter(NFL_rush, OffenseTeam == "PIT")
print(mean(NFL_pass$Yards))
print(mean(NFL_rush$Yards))

```

```{r}
Center = filter(NFL_rush, RushDirection == "CENTER")
LeftEnd = filter(NFL_rush, RushDirection == "LEFT END")
LeftGuard = filter(NFL_rush, RushDirection == "LEFT GUARD")
LeftTackle = filter(NFL_rush, RushDirection == "LEFT TACKLE")
RightEnd = filter(NFL_rush, RushDirection == "RIGHT END")
RightGuard = filter(NFL_rush, RushDirection == "RIGHT GUARD")
RightTackle = filter(NFL_rush, RushDirection == "RIGHT TACKLE")
print(mean(Center$Yards))
print(mean(LeftEnd$Yards))
print(mean(LeftGuard$Yards))
print(mean(LeftTackle$Yards))
print(mean(RightEnd$Yards))
print(mean(RightGuard$Yards))
print(mean(RightTackle$Yards))
```

```{r}
ShortLeft = filter(NFL_pass, PassType == "SHORT LEFT")
ShortRight = filter(NFL_pass, PassType == "SHORT RIGHT")
ShortMid = filter(NFL_pass, PassType == "SHORT MIDDLE")
print(mean(ShortLeft$Yards))
print(mean(ShortRight$Yards))
print(mean(ShortMid$Yards))
```




```{r}

library(ggplot2)
ggplot(data = NFL_rush, aes(x = Quarter, y = Yards, color = RushDirection)) + geom_point() + ggtitle("2022 NFL Rush Plays on Opponent 10 Yard Line and In" ) + theme(plot.title = element_text(hjust = 0.5))

ggplot(data = NFL_pass, aes(x = Quarter, y = Yards, color = PassType)) + geom_point() + ggtitle("2022 NFL Pass Plays on Opponent 10 Yard Line and In" ) + theme(plot.title = element_text(hjust = 0.5))


```


```{r}

library(ggplot2)
ggplot(data = NFL_Chiefsrush, aes(x = Quarter, y = Yards, color = RushDirection)) + geom_point() + ggtitle("2022 KC Rush Plays on Opponent 10 Yard Line and In" ) + theme(plot.title = element_text(hjust = 0.5))

ggplot(data = NFL_Chiefspass, aes(x = Quarter, y = Yards, color = PassType)) + geom_point() + ggtitle("2022 KC Pass Plays on Opponent 10 Yard Line and In" ) + theme(plot.title = element_text(hjust = 0.5))

ggplot(data = NFL_Steelersrush, aes(x = Quarter, y = Yards, color = RushDirection)) + geom_point() + ggtitle("2022 PIT Rush Plays on Opponent 10 Yard Line and In" ) + theme(plot.title = element_text(hjust = 0.5))

ggplot(data = NFL_Steelerspass, aes(x = Quarter, y = Yards, color = PassType)) + geom_point() + ggtitle("2022 PIT Pass Plays on Opponent 10 Yard Line and In" ) + theme(plot.title = element_text(hjust = 0.5))


```



```{r}

NFL_rush$RushDirection = as.factor(NFL_rush$RushDirection)

encoded_data = model.matrix(~ RushDirection - 1, data = NFL_rush)
combined_data = data.frame(encoded_data, Yards = NFL_rush$Yards)

rushmodel = lm(Yards ~ ., data = combined_data)


summary(rushmodel)

```



```{r}

NFL_pass$PassType = as.factor(NFL_pass$PassType)

encoded_data = model.matrix(~ PassType - 1, data = NFL_pass)
combined_data = data.frame(encoded_data, Yards = NFL_pass$Yards)

passmodel = lm(Yards ~ ., data = combined_data)


summary(passmodel)

```

```{r}

logisticNFL1 = glm(IsTouchdown ~ IsPass, family = binomial(), data = NFL_one)
logisticNFL2 = glm(IsTouchdown ~ IsRush, family = binomial(), data = NFL_one)
summary(logisticNFL1)
summary(logisticNFL2)
```



